---
title: "Leptograpsus crab morphology manuscript"
author: "Abi Brown"
Date: "2024-09-15"
output: pdf_document
header-includes:
  \usepackage{float}
 \floatplacement{figure}{H}
bibliography: 
  - references.bib
  - grateful-refs.bib
csl: "forest-science.csl"
---

##Introduction

*Leptograpsus variegatus* is a species of small shore crab found throughout the Southern Pacific, including on the shores of western Australia where the crabs dataset was collected [@campbell1974]. These crabs play an important role in the flow of nutrients through mangrove ecosystems, where they break down mangrove leaves and provide particulate organic matter to many species of detritivores [@camilleri1992].

#Data and analysis

Data for this project was originally collected and published in @campbell1974.

![Histograms showing distributions of frontal lobe (FL), carapace length (CL), and body depth (BD).](images/histograms.pdf)

```{r echo=FALSE}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```

##Results

Results of the PCA demonstrated some distinction between blue and orange morphs along the first principal component, and a larger distinction between males and females along the second axis of variation as seen in the following code chunk.

```{r crabsPCA, echo=TRUE}
library(vegan)
library(ggplot2)
library(ggordiplots)
crabs <- read.csv('../00_rawdata/crabs_measurements.csv')
#pca using vegan's rda function
pca_crabs <- rda(crabs[5:9], scale = TRUE)
#creating a species+sex column for visualization in the plot
crabs$spsex <- paste0(crabs$sp, crabs$sex)
#using ggordiplot. Ellipses are 1 standard deviation about the centroid.
gg_ordiplot(pca_crabs, groups = crabs$spsex, pt.size = 1, kind = 'sd')
```

##References
